---
title: "PCC conversion"
output: html_document
author: "Yefeng Yang and Shinichi Nakagawa"
date: '2023-02-24'
---

# Packages and function

```{r}
library(here)
library(dplyr)
library(readr)

# calculate t according to its definition in two-sample test
tval <- function(n1, n2, sd1, sd2, m1, m2) { # function to calculate independent samples t-test
  sdp <- ((n1 - 1) * sd1^2 + (n2 - 1) * sd2^2) / (n1 + n2 - 2) # calculate pooled SD
  tval <- (m1 - m2) / (sdp * sqrt(1 / n1 + 1 / n2)) 
  return(tval) 
}
```


# lnRR and SMD datasets with raw data (i.e., mean, sd, n)

Raw data mean the original authors used lnRR or SMD as effect size measure in their publications.

For this type of data, t and df can be calculated from raw data. Then you can use the two quantities to calculate pcc, correct and wrong SE(pcc) as you like.

```{r}
# lnRR 
dat_r_lnRR <- suppressMessages(list.files(here("data/lnRR/des_stat"), pattern = "*.csv", full.names = TRUE) %>% lapply(read_csv))

# calculate t and df
dat_r_lnRR_list <- list(NULL)
for (i in 1:length(dat_r_lnRR)) {
  dat_r_lnRR_list[i] <- dplyr::select(dat_r_lnRR[[i]], study_ID, T_mean, C_mean, T_sd, C_sd, T_n, C_n) %>% list() # keep useful variables
  dat_r_lnRR_list[i] <- mutate(dat_r_lnRR_list[[i]], t = tval(n1=T_n, n2=C_n, sd1=T_sd, sd2=C_sd, m1=T_mean, m2=C_mean), # calculate t according to its definition in independent t test
                                  df = T_n + C_n - 2)  %>% list()}

saveRDS(dat_r_lnRR_list, file = "data/converted/dat_r_lnRR_list.RDS")


# SMD 
dat_r_SMD <- suppressMessages(list.files(here("data/SMD/des_stat"), pattern = "*.csv", full.names = TRUE) %>% lapply(read_csv))

# calculate t and df
dat_r_SMD_list <- list(NULL)
for (i in 1:length(dat_r_SMD)) {
  dat_r_SMD_list[i] <- dplyr::select(dat_r_SMD[[i]], study_ID, T_mean, C_mean, T_sd, C_sd, T_n, C_n) %>% list() # keep useful variables
  dat_r_SMD_list[i] <- mutate(dat_r_SMD_list[[i]], t = tval(n1=T_n, n2=C_n, sd1=T_sd, sd2=C_sd, m1=T_mean, m2=C_mean), # calculate t according to its definition in independent t test
                                  df = T_n + C_n - 2)  %>% list()}
saveRDS(dat_r_SMD_list, file = "data/converted/dat_r_SMD_list.RDS")

```


# SMD datasets with pre-calculated effect sizes

Pre-calculated effect sizes data mean the original authors provided only effect size estimates and corresponding sampling variances in their publications. 

For this type of data, only SMD is possible to be converted into pcc and df. Note that pcc and df are approximated assuming that the equal designs (n1 = n2) were used in the studies. With the two quantities, you can calculate calculate correct and wrong SE(pcc) as you like.

```{r}
dat_p_SMD <- suppressMessages(list.files(here("data/SMD"), pattern = "*.csv", full.names = TRUE) %>% lapply(read_csv))

# calculate t and df
dat_p_SMD_list <- list(NULL)
for (i in 1:length(dat_p_SMD)) {
  dat_p_SMD_list[i] <- dplyr::select(dat_p_SMD[[i]], study_ID, es, var) %>% list() # keep useful variables
  dat_p_SMD_list[i] <- mutate(dat_p_SMD_list[[i]], pcc = es / sqrt(es^2 + 4)) %>% list() # convert d into pcc 
  dat_p_SMD_list[i] <- mutate(dat_p_SMD_list[[i]], n = round((8 + es^2)/(4*var)), # n = sample size of each group within each study
                               df = 2*n - 2) %>% list() # degrees of freedom
  }

saveRDS(dat_p_SMD_list, file = "data/converted/dat_p_SMD_list.RDS")


```


# Zr datasets

Zr data mean the original authors used Zr as effect measure in their publications. 

For this type of data, Zr can be back-transformed into pcc and df can be calculated based on var(Zr). Note that pcc and df from Zr are exactly calculated rather than approximated. With the two quantities, you can calculate calculate correct and wrong SE(pcc) as you like.

```{r}
dat_p_Zr <- suppressMessages(list.files(here("data/Zr"), pattern = "*.csv", full.names = TRUE) %>% lapply(read_csv))

# calculate t and df
dat_p_Zr_list <- list(NULL)
for (i in 1:length(dat_p_Zr)) {
  dat_p_Zr_list[i] <- dplyr::select(dat_p_Zr[[i]], study_ID, es, var) %>% list() # keep useful variables
  dat_p_Zr_list[i] <- mutate(dat_p_Zr_list[[i]], pcc = (exp(2*es) - 1)/(1 + exp(2*es))) %>% list() # convert Zr into pcc 
  dat_p_Zr_list[i] <- mutate(dat_p_Zr_list[[i]], n = round(1/var + 3), # sample size of each study
                          df = n - 2) %>% list() # degrees of freedom
  }

saveRDS(dat_p_Zr_list, file = "data/converted/dat_p_Zr_list.RDS")

```

